---
title: "Jenny_EDA"
output: html_document
date: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r load-packages, message=FALSE}
library(tidyverse)
library(broom)
library(patchwork)
```

```{r data}
df_1 <- read_csv("comma-survey.csv")
# df_1
colnames(df_1)

colSums(is.na(df_1)) # how did we deal with these before?
```

# Introduction
## Background

The importance of the use of the Oxford comma. (to make sentences less ambiguous)
-times when the lack of an Oxford comma failed us: 2018 Dairy Farmer and Maine.
-arguments against the use of the Oxford comma.

## Dataset Description and Variables of Interest
## Research Question
## EDA
# Methodology
## Model Selection
## Model Specification
## Model Diagnostics
# Results
# Discussion
## Investigative Conclusions
## Limitations and Future Work
- It's not a matter of grammar; it's a matter of clarity - Perlman. Would be
interesting to have data on whether people know the reasoning behind the Oxford comma.

# Summary

```{r}
### rename column names.
df_2 <- df_1 %>% 
  rename(comma_usage = "In your opinion, which sentence is more gramatically correct?",
         heard_comma = "Prior to reading about it above, had you heard of the serial (or Oxford) comma?",
         care_comma = "How much, if at all, do you care about the use (or lack thereof) of the serial (or Oxford) comma in grammar?",
         plural_usage = "How would you write the following sentence?",
         consider_plural = "When faced with using the word \"data\", have you ever spent time considering if the word was a singular or plural noun?",
         debate_plural = "How much, if at all, do you care about the debate over the use of the word \"data\" as a singluar or plural noun?",
         Grammar_isImportant = "In your opinion, how important or unimportant is proper use of grammar?"
         )
```

### Look at contingency table
```{r}
# create plots of 
df_2 %>% 
  group_by(comma_usage, care_comma) %>% 
  summarize(n=n())%>%
  mutate(proportion=n/sum(n))

df_2 %>% 
  group_by(comma_usage, heard_comma) %>% 
  summarize(n=n())%>%
  mutate(proportion=n/sum(n))

df_2 %>% 
  group_by(comma_usage, Grammar_isImportant) %>% 
  summarize(n=n())%>%
  mutate(proportion=n/sum(n))
```
Interestingly, among the people who have heard of the Oxford comma, 47.54% of these people chose not to use it. Even more interesting, among the people who self-reported "caring a lot" about the Oxford comma, 13.73% chose not to use it. Finally, among the people who thought that grammar was "Very Important", 62% did not use the Oxford comma.


# Research Hypothesis:
(cstudy1)
Control for the demographic features that are associated with use of the Oxford comma? (perhaps use a mixed effects model for different U.S. regions and justify using the 538 article.)
What region of the U.S. cares the most about Oxford comma?
Is the belief that grammar is important (variables: debate, consider, care), associated with correct usage?

