---
title: "Jenny_EDA"
output: html_document
date: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r load-packages, message=FALSE}
library(tidyverse)
library(broom)
library(forcats)
library(patchwork)
```

```{r data}
df_1 <- read_csv("comma-survey.csv")
# df_1
colnames(df_1)

colSums(is.na(df_1)) # how did we deal with these before?
```

# Introduction
## Background

The importance of the use of the Oxford comma. (to make sentences less ambiguous)
-times when the lack of an Oxford comma failed us: 2018 Dairy Farmer and Maine.
-arguments against the use of the Oxford comma.

## Dataset Description and Variables of Interest
## Research Question
## EDA


# Methodology
## Model Selection
## Model Specification
## Model Diagnostics


# Results

# Discussion
## Investigative Conclusions
## Limitations and Future Work
- It's not a matter of grammar; it's a matter of clarity - Perlman. Would be
interesting to have data on whether people know the reasoning behind the Oxford comma.

# Summary

```{r}
### rename column names.
df_2 <- df_1 %>% 
  rename(comma_usage = "In your opinion, which sentence is more gramatically correct?",
         heard_comma = "Prior to reading about it above, had you heard of the serial (or Oxford) comma?",
         care_comma = "How much, if at all, do you care about the use (or lack thereof) of the serial (or Oxford) comma in grammar?",
         plural_usage = "How would you write the following sentence?",
         consider_plural = "When faced with using the word \"data\", have you ever spent time considering if the word was a singular or plural noun?",
         debate_plural = "How much, if at all, do you care about the debate over the use of the word \"data\" as a singluar or plural noun?",
         Grammar_isImportant = "In your opinion, how important or unimportant is proper use of grammar?"
         )

df_3 <- df_2 %>% 
  mutate(care_comma = fct_relevel(care_comma, 
            "Not at all","Not much",
                               "Some",
                               "A lot")) %>% 
  mutate(Grammar_isImportant = fct_relevel(Grammar_isImportant, 
            "Somewhat unimportant",
            "Neither important nor unimportant (neutral)",
                               "Somewhat important",
                               "Very important"))

```

### Look at contingency table
```{r}
# create plots of 
df_3 %>% 
  group_by(comma_usage, care_comma) %>% 
  summarize(n=n())%>%
  mutate(proportion=n/sum(n))

df_3 %>% 
  group_by(comma_usage, heard_comma) %>% 
  summarize(n=n())%>%
  mutate(proportion=n/sum(n))

df_3 %>% 
  group_by(comma_usage, Grammar_isImportant) %>% 
  summarize(n=n())%>%
  mutate(proportion=n/sum(n))
```

```{r}
ggplot(data = df_3) +
  geom_count(mapping = aes(x=comma_usage, y=care_comma))+ 
  scale_x_discrete(name ="Oxford Comma", 
                   labels=c("omitted", 
                            "used"
                  ))+
  labs(x = "Oxford Comma", y = "Care about Comma")+
  theme_bw()

# Heat map.
df_3 %>% 
  count(comma_usage, care_comma) %>%  
  ggplot(mapping = aes(x = comma_usage, y = care_comma)) +
    geom_tile(mapping = aes(fill = n)) +
  scale_x_discrete(name ="Oxford Comma", 
                   labels=c("omitted", 
                            "used"
                  ))+labs(x = "Oxford Comma", y = "Care about Comma")

ggplot(data = df_3) +
  geom_count(mapping = aes(x=comma_usage, y=Grammar_isImportant))+ 
  scale_x_discrete(name ="Oxford Comma", 
                   labels=c("omitted", 
                            "used"
                  ))+
  labs(x = "Oxford Comma", y = "Grammar Is Important")+
  theme_bw()

# Heat map.
df_3 %>% 
  count(comma_usage, Grammar_isImportant) %>%  
  ggplot(mapping = aes(x = comma_usage, y = Grammar_isImportant)) +
  geom_tile(mapping = aes(fill = n)) +
    scale_x_discrete(name ="Oxford Comma", 
                   labels=c("omitted", 
                            "used"
                  ))+
  labs(x = "Oxford Comma", y = "Grammar Is Important")
```

Interestingly, among the people who have heard of the Oxford comma, 47.54% of these people chose not to use it. Even more interesting, among the people who self-reported "caring a lot" about the Oxford comma, 13.73% chose not to use it. Finally, among the people who thought that grammar was "Very Important", 62% did not use the Oxford comma.

The distribution between people who omitted and used the comma were very similar.


# Research Hypothesis:
(cstudy1)
Control for the demographic features that are associated with use of the Oxford comma? (perhaps use a mixed effects model for different U.S. regions and justify using the 538 article.)

Is the belief that grammar is important (variables: debate, consider, care), associated with correct usage?

